SI Appendix

### 2024-05-09 manuscript/revision2/intermediate_files/response_to_reviews.qmd

---
title: "Response_to_Reviews"
format: pdf
editor: visual
---

**Reviewer comment**

Mash: A lot of the results interpretation is based on interpreting the SNP loadings on the specified covariance matrices. This is a secondary use of mash (the primary being the refinement of effect sizes), and while they do interpret these somewhat in the mash paper, these loadings need to be interpreted with caution. If these covariance matrices are similar, mash will somewhat arbitrarily assign weight to any of the matrices because all lead to the same fit to the effect-size data. The mash algorithm doesn't give the full posterior distribution on the loadings so you can't check for posterior correlations there. It seems likely to me that the hypothesis-driven and data-driven covariance matrices are somewhat correlated here, and the correlations may differ between green-up and flowering because of the better correlation between the environmental metrics and flowering than green-up. In the mash paper, they used cross-validation with the likelihood in the testing set as the evaluation metric to compare models. It might be safer to try dropping specific covariance matrices and comparing the model performance in a held-out testing set to evaluate the importance of the hypothesis-derived covariance matrices.

@@

likelihood = -inf

most_likely_matrix = ''

for matrix in matrices:

```         
matrix_likelihood = mash(effects, std.errs, matrix)\[\'likelihood\'\] #obtain model likelihood

@@

lrt_pvalue = 0

iteration = 2

while lrt_pvalue \< 0.05:

```         
likelihood = -inf
```

most_likely_matrix = ''

```         
for matrix in matrices:

@@

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\

**Methods**

In order to better understand the behavior of the mash shrinkage algorithm in the presence of correlated covariance matrices, we implemented a "greedy" mash algorithm. Specifically, given n covariance matrices for each phenotype and corresponding effect sizes and standard errors across the k contexts of interest, we fit a set of mash models with each of the n matrices separately. We then identify the model with the maximum log likelihood estimates and select the corresponding matrix, leaving n-1 covariance matrices. We then fit a new set of mash models where the selected covariance matrix is paired with each of the remaining n-1 covariance matrices. The most likely pair of covariance matrices are then selected and the process is repeated for all possible n-2 matrix triplets. This process is repeated until one of two stop conditions are met: (i) all covariance matrices have been added to the mash model, resulting in a final model with all of the original n covariance matrices are included or (ii) a likelihood ratio test with one degree of freedom between the likelihood of the current model and the model from the previous iteration has a p-value \> 0.05. The result is a set of log likelihood estimates for each model and the stepwise 'path' of covariance matrices that best explain the observed effect sizes and standard errors observed for the phenotype of interest.

### 2024-05-09 manuscript/revision2/intermediate_files/response_to_reviews.tex

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Response\_to\_Reviews},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Response\_to\_Reviews}
\author{}
\date{}

\begin{document}
\maketitle

\textbf{Reviewer comment}

Mash: A lot of the results interpretation is based on interpreting the
SNP loadings on the specified covariance matrices. This is a secondary
use of mash (the primary being the refinement of effect sizes), and
while they do interpret these somewhat in the mash paper, these loadings
need to be interpreted with caution. If these covariance matrices are
similar, mash will somewhat arbitrarily assign weight to any of the
matrices because all lead to the same fit to the effect-size data. The
mash algorithm doesn't give the full posterior distribution on the
loadings so you can't check for posterior correlations there. It seems
likely to me that the hypothesis-driven and data-driven covariance
matrices are somewhat correlated here, and the correlations may differ
between green-up and flowering because of the better correlation between
the environmental metrics and flowering than green-up. In the mash
paper, they used cross-validation with the likelihood in the testing set
as the evaluation metric to compare models. It might be safer to try
dropping specific covariance matrices and comparing the model
performance in a held-out testing set to evaluate the importance of the
hypothesis-derived covariance matrices.

@@

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\hfill\break

\textbf{Methods}

In order to better understand the behavior of the mash shrinkage
algorithm in the presence of correlated covariance matrices, we
implemented a ``greedy'' mash algorithm. Specifically, given n
covariance matrices for each phenotype and corresponding effect sizes
and standard errors across the k contexts of interest, we fit a set of
mash models with each of the n matrices separately. We then identify the
model with the maximum log likelihood estimates and select the
corresponding matrix, leaving n-1 covariance matrices. We then fit a new
set of mash models where the selected covariance matrix is paired with
each of the remaining n-1 covariance matrices. The most likely pair of
covariance matrices are then selected and the process is repeated for
all possible n-2 matrix triplets. This process is repeated until one of
two stop conditions are met: (i) all covariance matrices have been added
to the mash model, resulting in a final model with all of the original n
covariance matrices are included or (ii) a likelihood ratio test with
one degree of freedom between the likelihood of the current model and
the model from the previous iteration has a p-value \textgreater{} 0.05.
The result is a set of log likelihood estimates for each model and the
stepwise `path' of covariance matrices that best explain the observed
effect sizes and standard errors observed for the phenotype of
interest.~

### 2024-05-09 manuscript/revision2/intermediate_files/SI_Appendix.qmd
manuscript: add response to reviews folder within manuscript/revision2/ and move intermediate files to intermediate_files folder

@@

+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

|           | where \'**n**\' is in {5, 10, 18} |                                                                                                                                      |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

|           | where \'**n**\' is in {1, 3, 5}   |                                                                                                                                      |
+-----------+-----------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

	We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1). We defined flowering as functions of eight variables: \'flowering date\' (day of the year), cumulative growing degree days (GDD) between green-up and flowering dates (\'flowering GDD\'), day length on the flowering date (\'flowering daylength\'), change in daylength relative to the previous day on the day of flowering (\'flowering daylength change\'), and four measures of cumulative rainfall: cumulative rainfall between green-up and flowering, and in the one day, three days, and five days before flowering. Cumulative GDD was calculated as GDD = GRFLmaxTmean−Tbase,0, where Tmean is the daily average temperature, defined as (Tmax+Tmin)/2, Tbase is the base temperature of 12 °C for switchgrass, Tmax is the maximum daily temperature, Tmin is the minimum daily temperature, GR is green-up date, and FL is flowering date (Kiniry et al. 2005; Behrman et al. 2013).  On a specific day, if Tmean is less than Tbase, the GDD for that day is 0; if Tmean is bigger than Tbase, the GDD for that day is the difference between Tmean and Tbase. We also defined green-up as functions of seven variables: \'green-up date\' (day of the year), cumulative GDD for the five, ten, or eighteen days prior to green-up, and the average air temperature for the five, ten, or eighteen days prior to green-up.

@@

	We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1). These covariance matrices were derived from the same set of phenotypes as environmental functions of green-up and flowering, and represent the correlations between genotypes for these phenotypes across our common gardens. For the diagonal of these matrices, we used the coefficient of variation in these phenotypes within the subpopulation at that garden. Code used to generate these matrices can be found at  [git\@github.com](mailto:git@github.com):Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd. We used distinct sets of user-specified, hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups. If a hypothesis-based matrix does not capture a common pattern of SNP effects, mash assigns small mixture proportions onto this matrix using maximum likelihood, giving that covariance matrix a low posterior weight summed across all SNP effects, which ultimately causes little to no change in the mash model. Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.

@@

	In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al. 2009). Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both). We used ASReml (VSN International) to specify mixed models of the form:

@@

	These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term. This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).

@@

	In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models. We used mixed models of the form:

@@

	These models were run for each phenotype as a function of environmental cues. This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).

@@

	To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al. 2019). Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions. To this end, mash allows the user to include hypothesis-based covariance matrices, specifies some \'canonical\' covariance matrices with simple patterns of effects, and also generates \'data-driven\' covariance matrices from patterns of effects in the data. Then, mash assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood. Finally, mash uses Bayes\' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions. For example, a SNP affecting flowering may have a high mixture proportion, or weight, on a covariance matrix created from a specific environmental cue. In that case, we can infer that the effect of that SNP on flowering is caused by a response to or interaction with that environmental cue. These user-specified and data-driven covariance matrices are an important advantage mash offers for studying patterns of GxE: the user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP, while the data-driven covariance matrices allow exploration of additional unexplained patterns of covariation.

	We first conducted univariate GWAS at each common garden for green-up and flowering date, then analyzed the allelic effects for the top 19K unlinked SNPs per univariate GWAS using mash, as in Lovell et al. (2021). To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (Korte and Farlow 2013), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together (hereafter \'Both\' subpopulations). Only SNPs with ≤ 20% missing data and minor allele frequencies \> 0.05 were used in univariate GWAS, resulting in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations. Then, we created a set of \'random\' markers to account for correlation in effects using a set of 19K relatively unlinked (r2 \< 0.2) markers. Next, we created a set of \'strong\' markers by selecting 19K SNPs with the largest -log10(p-value) in any condition in the univariate GWAS, out of all markers unlinked at (r2 \< 0.2), which represented the SNP with the strongest effect in the 19K LD blocks with the largest univariate effects. We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups.

@@

	To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously (Milano, Lowry, and Juenger 2016). The parents of this cross were DAC, an early flowering Midwest individual, VS16, a late flowering Midwest individual, AP13, an early flowering Gulf individual, and WBC, a late flowering Gulf individual. We made F1 crosses of the two early flowering genotypes, AP13xDAC, and the two late flowering genotypes, WBCxVS16. We then clonally propagated and planted the four parents, the two F1 genotypes (AP13xDAC, and VS16xWBC), and 801 F2 genotypes at eight field sites in May-July of 2015. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used here.

	Details on the genetic map construction, map polishing and fine-scale reordering can be accessed on <https://datadryad.org/stash/dataset/doi:10.5061/dryad.ghx3ffbjv>. QTL mapping was conducted with R/qtl2 (Broman et al. 2019). We performed a genome scan with a linear mixed model that accounts for the relationships among individuals and for environmental covariates (i.e., field sites). The full model can be expressed as: phenotype = µ + QTL + E + QTLxE + kinship + e, where µ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTLxE is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the \'scan1\' function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using \'scan1perm\' function. The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2.

>>>>

## Manuscript Length {.unnumbered}

A standard 6-page article is approximately 4,000 words, 50 references, and 4 medium-size graphical elements (i.e., figures and tables). The preferred length of articles remains at 6 pages, but PNAS will allow articles up to a maximum of 12 pages.

### Digital Figures {.unnumbered}

EPS, high-resolution PDF, and PowerPoint are preferred formats for figures that will be used in the main manuscript. Authors may submit PRC or U3D files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Images must be provided at final size, preferably 1 column width (8.7cm). Figures wider than 1 column should be sized to 11.4cm or 17.8cm wide. Numbers, letters, and symbols should be no smaller than 6 points (2mm) and no larger than 12 points (6mm) after reduction and must be consistent.

Figures and tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.

To insert a figure wider than one column, please use the `figure*` environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide. Use `SCfigure*` environment for a wide figure with side legends, such as @fig-side.

In Quarto, we can do these by setting the `fig-env` command to `figure*` or `SCfigure*`

```{r}
#| label: fig-side
#| fig-cap: "This legend would be placed at the side of the figure, rather than below it."
#| fig-env: "SCfigure*"
#| echo: false
# tell it the options as comments with a | and a space, as above.
# set echo: false to avoid printing this text
knitr::include_graphics('figs/frog.pdf')
```

## Supporting Information Appendix (SI)

Authors should submit SI as a single separate SI Appendix PDF file, combining all text, figures, tables, movie legends, and SI references. SI will be published as provided by the authors; it will not be edited or composed. Additional details can be found in the [PNAS Author Center](https://www.pnas.org/authors/submitting-your-manuscript#manuscript-formatting-guidelines). The PNAS Overleaf SI template can be found [here](https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd). Refer to the SI Appendix in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc.Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.

## Data Archival {.unnumbered}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the [Information for Authors](https://www.pnas.org/author-center/editorial-and-journal-policies#materials-and-data-availability), is acceptable.

@@

+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

|           | where '**n**' is in {5, 10, 18} |                                                                                                                                      |
+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

|           | where '**n**' is in {1, 3, 5}   |                                                                                                                                      |
+-----------+---------------------------------+--------------------------------------------------------------------------------------------------------------------------------------+

@@

```         
We scored green-up and flowering by day of the year, then linked these dates to multiple environmental factors (Table S1). We defined flowering as functions of eight variables: \'flowering date\' (day of the year), cumulative growing degree days (GDD) between green-up and flowering dates (\'flowering GDD\'), day length on the flowering date (\'flowering daylength\'), change in daylength relative to the previous day on the day of flowering (\'flowering daylength change\'), and four measures of cumulative rainfall: cumulative rainfall between green-up and flowering, and in the one day, three days, and five days before flowering. Cumulative GDD was calculated as GDD = GRFLmaxTmean−Tbase,0, where Tmean is the daily average temperature, defined as (Tmax+Tmin)/2, Tbase is the base temperature of 12 °C for switchgrass, Tmax is the maximum daily temperature, Tmin is the minimum daily temperature, GR is green-up date, and FL is flowering date (Kiniry et al. 2005; Behrman et al. 2013).  On a specific day, if Tmean is less than Tbase, the GDD for that day is 0; if Tmean is bigger than Tbase, the GDD for that day is the difference between Tmean and Tbase. We also defined green-up as functions of seven variables: \'green-up date\' (day of the year), cumulative GDD for the five, ten, or eighteen days prior to green-up, and the average air temperature for the five, ten, or eighteen days prior to green-up.

```

@@

```         
We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three subpopulations (Table S1). These covariance matrices were derived from the same set of phenotypes as environmental functions of green-up and flowering, and represent the correlations between genotypes for these phenotypes across our common gardens. For the diagonal of these matrices, we used the coefficient of variation in these phenotypes within the subpopulation at that garden. Code used to generate these matrices can be found at  [git\@github.com](mailto:git@github.com):Alice-MacQueen/pvdiv-phenology-gxe.git/R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd. We used distinct sets of user-specified, hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups. If a hypothesis-based matrix does not capture a common pattern of SNP effects, mash assigns small mixture proportions onto this matrix using maximum likelihood, giving that covariance matrix a low posterior weight summed across all SNP effects, which ultimately causes little to no change in the mash model. Alternatively, SNPs with high mixture proportions on particular environmental covariance matrices and large Bayes factors, which summarize the overall significance of a non-zero effect, represent small genomic intervals with strong evidence for a phenotypic effect correlated with an environmental driver.
```

@@

```         
In the diversity panel, we determined narrow-sense heritabilities (h2) for green-up and flowering dates at single gardens using genomic relationship matrices calculated using the van Raden method (VanRaden et al. 2009). Genomic relationship matrices were calculated within each subpopulation (Midwest and Gulf) and for both genetic subpopulations (Both). We used ASReml (VSN International) to specify mixed models of the form:
```

@@

```         
These models were run for each of the eight gardens, and across all gardens by adding an additional environmental effect of site without an interaction term. This resulted in 54 models: 3 sets of populations (the Gulf, Midwest, and Both subpopulations) for 9 garden sets (all eight gardens separately, and all eight gardens together) and two phenotypes (green-up date and flowering date).
```

@@

```         
In the diversity panel, to evaluate our environmental cues as genetic triggers of flowering, we defined green-up and flowering for individual genotypes as functions of seven and eight environmental cues, respectively, then used variance components analysis to partition variance attributed to genetic effects (G), genotype by environment interactions (GxE), environmental effects (E), and error for these phenology-related traits using linear mixed models. We used mixed models of the form:
```

@@

```         
These models were run for each phenotype as a function of environmental cues. This resulted in 135 models: three sets of populations (the Gulf, Midwest, and Both subpopulations) for three garden sets (All, North, and Texas gardens) and 15 phenotypes (seven green-up functions, and eight flowering functions).
```

@@

```         
To evaluate the prevalence and kinds of covariance patterns of SNP effects across our eight common gardens, we used multivariate adaptive shrinkage (mash) on SNP effect estimates from the diversity panel (Urbut et al. 2019). Mash is a statistical method that allows estimation and comparison of many effects jointly across many different conditions; it improves on previous methods by allowing arbitrary correlations in effect sizes among conditions. To this end, mash allows the user to include hypothesis-based covariance matrices, specifies some \'canonical\' covariance matrices with simple patterns of effects, and also generates \'data-driven\' covariance matrices from patterns of effects in the data. Then, mash assigns mixture proportions for each SNP onto each provided covariance matrix using maximum likelihood. Finally, mash uses Bayes\' theorem to shrink effects for each SNP towards the set of covariance matrices in accordance to their mixture proportions. For example, a SNP affecting flowering may have a high mixture proportion, or weight, on a covariance matrix created from a specific environmental cue. In that case, we can infer that the effect of that SNP on flowering is caused by a response to or interaction with that environmental cue. These user-specified and data-driven covariance matrices are an important advantage mash offers for studying patterns of GxE: the user-specified covariance matrices allow hypothesis testing of specific environmental drivers for each SNP, while the data-driven covariance matrices allow exploration of additional unexplained patterns of covariation.

We first conducted univariate GWAS at each common garden for green-up and flowering date, then analyzed the allelic effects for the top 19K unlinked SNPs per univariate GWAS using mash, as in Lovell et al. (2021). To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes (Korte and Farlow 2013), we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together (hereafter \'Both\' subpopulations). Only SNPs with ≤ 20% missing data and minor allele frequencies \> 0.05 were used in univariate GWAS, resulting in 8.8 million SNPs retained for the Midwest subpopulation, 10.3 million SNPs retained for the Gulf subpopulation, and 12.3 million SNPs retained for Both subpopulations. Then, we created a set of \'random\' markers to account for correlation in effects using a set of 19K relatively unlinked (r2 \< 0.2) markers. Next, we created a set of \'strong\' markers by selecting 19K SNPs with the largest -log10(p-value) in any condition in the univariate GWAS, out of all markers unlinked at (r2 \< 0.2), which represented the SNP with the strongest effect in the 19K LD blocks with the largest univariate effects. We used the \'pvdiv_standard_run()\' function of the switchgrassGWAS R package (https://github.com/Alice-MacQueen/switchgrassGWAS) to conduct GWAS on genotypes from the Midwest, Gulf, and Both subpopulations, used the \'pvdiv_bigsnp2mashr()\' function to convert univariate GWAS output to the matrices of effects and standard errors needed for mash, then used the \'mash_standard_run()\' function to conduct mash, again specifying distinct sets of hypothesis-based covariance matrices for green-up and flowering, but the same set of matrices for all genetic subgroups. 
```

@@

```         
To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously (Milano, Lowry, and Juenger 2016). The parents of this cross were DAC, an early flowering Midwest individual, VS16, a late flowering Midwest individual, AP13, an early flowering Gulf individual, and WBC, a late flowering Gulf individual. We made F1 crosses of the two early flowering genotypes, AP13xDAC, and the two late flowering genotypes, WBCxVS16. We then clonally propagated and planted the four parents, the two F1 genotypes (AP13xDAC, and VS16xWBC), and 801 F2 genotypes at eight field sites in May-July of 2015. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used here.

Details on the genetic map construction, map polishing and fine-scale reordering can be accessed on <https://datadryad.org/stash/dataset/doi:10.5061/dryad.ghx3ffbjv>. QTL mapping was conducted with R/qtl2 (Broman et al. 2019). We performed a genome scan with a linear mixed model that accounts for the relationships among individuals and for environmental covariates (i.e., field sites). The full model can be expressed as: phenotype = µ + QTL + E + QTLxE + kinship + e, where µ is the population mean, QTL is the marker genetic effect, E is the environmental effects (here, common garden), QTLxE is the interaction between marker genetic and environmental effects, kinship corresponds to the background polygenic variation, and e is the error term. The genome scan was accomplished with the \'scan1\' function. The statistical significance of the genome scan was established by performing a stratified (i.e., stratifying on common garden) permutation test (n=1000) using \'scan1perm\' function. The estimated QTL effect was obtained using \'scan1coef\' function in R/qtl2. 
```
